Host richness increases the occurrence but not the severity of bark beetle-induced tree mortality

### 2021-09-14 Code/Manuscript-SMOTE.Rmd

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = F,
	fig.align = "center",
	message = FALSE,
	warning = FALSE,
	dpi = 300,
	progress = FALSE,
	cache = FALSE
)

@@

```{r}
Fig.file <- here("Results/Figures/Figure-AgentbyHostcombo.jpg")
df$hostsFPshort<- df$hostsFPlong
df$hostsFPshort <- factor(df$hostsFPlong, levels=levels(df$hostsFPlong), labels=c("PICO", "PIEN", "ABLA", "PICO\n&\nPIEN", "PICO\n&\nABLA", "PIEN\n&\nABLA", "PICO,\nPIEN,\n&\nABLA"))
df$agentsFplongrev <- df$agentsFplong
df$agentsFplongrev <- factor(df$agentsFplongrev, levels=rev(levels( df$agentsFplong)))
jpeg(Fig.file, width=fig.width2,height=fig.width1.5,units="in", res=300)
ggplot(df[df$focalhost=="Any",], aes(x=hostsFPshort,fill=agentsFplongrev))+geom_bar()+theme(legend.title = element_blank())+xlab("hosts")+scale_fill_manual(values=p2col)
whatever <- dev.off()
```

### Given suitable stand conditions, is co-occurrence common?

Construct random forest models using synthetic minority oversampling technique (SMOTE)

#### MPB

```{r fit MPB RF model, eval=F}
response <- "agentMPB"
vars <- c("BA.PICO", "QMD.PICO", "propBA.PICO")
datBB.c <- df[df$focalhost=="Any",]

@@

```{r}
MPBpp <- read.csv(here("Results/RF-partialdependence-agentMPB-occurrence.csv"))
MPBpp$focalagent <- "MPB"
SBpp <- read.csv(here("Results/RF-partialdependence-agentSB-occurrence.csv"))
SBpp$focalagent <- "SB"
WBBBpp <- read.csv(here("Results/RF-partialdependence-agentWBBB-occurrence.csv"))
WBBBpp$focalagent <- "WBBB"

pBA1 + pBA2 +plot_layout(ncol=1)

ppropBA1 + ppropBA2 +plot_layout(ncol=1)

Fig.file <- here("Results/Figures/PartialDependence.jpg")
jpeg(Fig.file, width=fig.width2,height=fig.width1,units="in", res=300)
pBA1  + pQMD1 +theme(axis.title.y=element_blank(),axis.text.y=element_blank()) +ppropBA1 +theme(axis.title.y=element_blank(),axis.text.y=element_blank())+ pBA2 + pQMD2 +theme(axis.text.y=element_blank())  + ppropBA2 +theme(axis.text.y=element_blank()) + plot_layout(ncol=3, heights=c(2,1))
whatever <- dev.off()
```

```{r}
MPBpred <- read.csv(here("Results/RF-pred-agentMPB-occurrence.csv"))
SBpred <- read.csv(here("Results/RF-pred-agentSB-occurrence.csv"))
WBBBpred <- read.csv(here("Results/RF-pred-agentWBBB-occurrence.csv"))

@@

threshold <- 0.25
BBpred$MPBpred01 <- ifelse(BBpred$MPBpred>threshold, "MPB", "absent")
BBpred$SBpred01 <- ifelse(BBpred$SBpred>threshold, "SB", "absent")
BBpred$WBBBpred01 <- ifelse(BBpred$WBBBpred>threshold, "WBBB", "absent")

BBpred$agentspred <-  do.call(paste, c(BBpred[c("SBpred01", "MPBpred01", "WBBBpred01")],sep="-"))
BBpred$agentspred01 <- factor(BBpred$agentspred, levels=c("absent-absent-absent","absent-MPB-absent","SB-absent-absent", "SB-MPB-absent", "absent-absent-WBBB", "SB-absent-WBBB", "absent-MPB-WBBB", "SB-MPB-WBBB"), labels=c("none", "MPB", "SB", "MPB & SB", "WBBB", "SB & WBBB", "MPB & WBBB", "MPB, SB & WBBB"))

dfpred <- merge(BBpred, df[df$focalagent=="Any",], id.vars="PLT_CN", all=T)
```

```{r}
mpb <- as.data.frame(table(bbo.dat[bbo.dat$focalagent=="MPB","affected"], bbo.dat[bbo.dat$focalagent=="MPB","hostspred"]))
sb <- as.data.frame(table(bbo.dat[bbo.dat$focalagent=="SB","affected"], bbo.dat[bbo.dat$focalagent=="SB","hostspred"]))
wbbb <- as.data.frame(table(bbo.dat[bbo.dat$focalagent=="WBBB","affected"], bbo.dat[bbo.dat$focalagent=="WBBB","hostspred"]))
any<- as.data.frame(table(bbo.dat[bbo.dat$focalagent=="Any","affected"], bbo.dat[bbo.dat$focalagent=="Any","hostspred"]))

results <- NULL
for(j in levels(bbo.dat$hostspred)[5:8]){
  anyj <- any[any$Var2==j,]
  mpbj <- mpb[mpb$Var2==j,]
  sbj <- sb[sb$Var2==j,]
  wbbbj <- wbbb[wbbb$Var2==j,]

@@

  res <- data.frame(hostspred=NA, barkbeetle=NA, X=NA, p=NA, prop1=NA, prop2 =NA, lower=NA, upper=NA)
  res$hostspred <-j
  res$barkbeetle <- "MPB"
  res$X <- x$statistic
  res$p <- x$p.value
  res$prop1 <- x$estimate[1]
  res$prop2 <- x$estimate[2]
  res$lower <- x$conf.in[1]
  res$upper <- x$conf.in[2]
  x <-NULL
  results <- rbind(results, res)

@@

  res <- data.frame(hostspred=NA, barkbeetle=NA, X=NA, p=NA, prop1=NA, prop2 =NA, lower=NA, upper=NA)
  res$hostspred <-j
  res$barkbeetle <- "SB"
  res$X <- x$statistic
  res$p <- x$p.value
  res$prop1 <- x$estimate[1]
  res$prop2 <- x$estimate[2]
  res$lower <- x$conf.in[1]
  res$upper <- x$conf.in[2]
  x <-NULL
  results <- rbind(results, res)

@@

  res <- data.frame(hostspred=NA, barkbeetle=NA, X=NA, p=NA, prop1=NA, prop2 =NA, lower=NA, upper=NA)
  res$hostspred <-j
  res$barkbeetle <- "WBBB"
  res$X <- x$statistic
  res$p <- x$p.value
  res$prop1 <- x$estimate[1]
  res$prop2 <- x$estimate[2]
  res$lower <- x$conf.in[1]
  res$upper <- x$conf.in[2]
  x <-NULL
  results <- rbind(results, res)
}

@@

x <- bbo.dat[is.na(bbo.dat$hostspred)==FALSE,]
x <- x[x$hostspred %in% levels(x$hostspred)[5:8],]
x[x$hostspred == "PICO\n& PIEN" & x$focalagent=="WBBB","affected"]<- 0
x[x$hostspred == "PICO\n& ABLA" & x$focalagent=="SB","affected"]<- 0
x[x$hostspred == "PIEN\n& ABLA" & x$focalagent=="MPB","affected"]<- 0

xx <- x %>% group_by(hostspred,focalagent) %>% count(affected)
xxx <- x %>% group_by(hostspred) %>% count()

xxxx <- merge(xx, xxx, by="hostspred")
xxxx$prop <- xxxx$n.x/xxxx$n.y
xxxx <- xxxx[xxxx$affected=="affected",]
xxxx <- xxxx[,c(1,2,6)]
colnames(xxxx) <- c("hostspred", "barkbeetle", "prop")

resultsx <- merge(results, xxxx, id.vars=c("hostspred", "barkbeetle"), all=T)
resultsx$barkbeetle <- factor(resultsx$barkbeetle, levels=c("MPB", 'SB', "WBBB", "Any"), ordered=T)
resultsx <- resultsx[is.na(resultsx$prop)==FALSE,]
resultsx$hostspred <- factor(resultsx$hostspred, levels=levels(bbo.dat$hostspred)[5:8], ordered=T)
resultsx$pplot <- ifelse(resultsx$p<0.05, "*", "")
resultsx$pplot[is.na(resultsx$pplot) ==TRUE] <- ""

Fig.file <- here("Results/Figures/Figurey.jpg")
jpeg(Fig.file, width=fig.width2,height=fig.width1,units="in", res=300)
ggplot(resultsx, aes(x=barkbeetle,y=prop, fill=barkbeetle))+geom_bar(stat="identity")+theme(legend.title = element_blank())+geom_text(aes(label=pplot), vjust=-0.25,size=6)+facet_grid(.~hostspred, scales="free_x", space="free_x")+xlab("bark beetle species")+scale_fill_manual(values=c(p1col[1:3], gray(0.3)))+theme(legend.position = "none")+ylab("proportion of plots")
whatever <- dev.off()
```

```{r}
Fig.file <- here("Results/Figures/Figure-AgentbyHostcombo.jpg")
bbo.dat$hostsFPshort<- bbo.dat$hostsFPlong
bbo.dat$hostsFPshort <- factor(bbo.dat$hostsFPlong, levels=levels(bbo.dat$hostsFPlong), labels=c("PICO", "PIEN", "ABLA", "PICO\n&\nPIEN", "PICO\n&\nABLA", "PIEN\n&\nABLA", "PICO,\nPIEN,\n&\nABLA"))
bbo.dat$agentsFplongrev <- bbo.dat$agentsFplong
bbo.dat$agentsFplongrev <- factor(bbo.dat$agentsFplongrev, levels=rev(levels( bbo.dat$agentsFplong)))
#jpeg(Fig.file, width=fig.width2,height=fig.width1.5,units="in", res=300)
ggplot(bbo.dat[bbo.dat$focalhost=="Any",], aes(x=hostsFPshort,fill=agentsFplongrev))+geom_bar()+theme(legend.title = element_blank())+xlab("hosts")+scale_fill_manual(values=p2col)
#whatever <- dev.off()
```

### Is occurrence more likely in stands with multiple hosts?

### Is severity greater in stands with multiple hosts?

To determine if host diversity or identity influences the occurrence tree mortality attributed to bark beetles within a stand, we used simple Chi-square tests. Given a significant difference among groups, we then used pairwise Fisher's exact tests to determine pairwise differences, while accounting for multiple comparisons [@Benjamini_controlling_1995].

@@

```{r FigPresHosts, fig.cap = "The frequency of FIA plots inventoried by host identity and presence (dark) and absence (light) of bark beetle activity across the study area. ", out.width = "244.8pt"}
knitr::include_graphics( here("Results/Figures/Figure-PresencebyHostcombos.jpg"))
```

## Occurrence 
```{r FigPropAgentxHosts, fig.cap = "A caption for figure 1", out.width = "504pt", eval=F}
knitr::include_graphics( here("Results/Figures/Figure-PropAffectedByAgentXHosts.jpg"))
```

@@

@@

```{r FigOEACbyHI, fig.cap = "A caption for figure 2", out.width = "504pt"}
knitr::include_graphics(here("Results/Figures/Figure-ObsExpCo-ocurrenceAgentXHosts.jpg"))
```

## Ecological effects

@@

## Supplement
### Comparison of stand structure and composition, climate, and regional bark beetle activity  
```{r FigStandSC,  fig.cap = "A caption", out.width = "322.56pt"}
datBB.c <-df.sub[!df.sub$focalhost=="Any",]

@@

```{r FigPressure, fig.cap = "A caption", out.width = "322.56pt"}
## select plots where regional pressure can be estimated (> 5= plots)

RpressureSB.melt <- datBB.c[,c("hosts", "Rpressure.SB", "hostBA.SB", "nplots")]
colnames(RpressureSB.melt) <- c("hosts", "pressure", "hostBA", "nplots")
RpressureSB.melt$variable <- "SB"

RpressureMPB.melt <- datBB.c[, c("hosts", "Rpressure.MPB", "hostBA.MPB", "nplots")]
colnames(RpressureMPB.melt) <- c("hosts", "pressure", "hostBA", "nplots")
RpressureMPB.melt$variable <- "MPB"

RpressureWBBB.melt <- datBB.c[, c("hosts", "Rpressure.WBBB", "hostBA.WBBB", "nplots")]
colnames(RpressureWBBB.melt) <- c("hosts", "pressure", "hostBA", "nplots")
RpressureWBBB.melt$variable <- "WBBB"

Rpressure.melt <- rbind(RpressureMPB.melt,RpressureSB.melt)
Rpressure.melt <- rbind(Rpressure.melt,RpressureWBBB.melt)

Rpressure.melt$variable <- factor(Rpressure.melt$variable, levels=c("MPB","SB",  "WBBB"), labels=c("MPB","SB",  "WBBB"), ordered=T)
Rpressure.melt$propRegional <- Rpressure.melt$pressure /Rpressure.melt$hostBA
Rpressure.melt$host <- factor(Rpressure.melt$variable, levels=c("MPB","SB", "WBBB"), labels=c("PICO", "PIEN", "ABLA"), ordered=T)

ggpropRegional <- ggplot(Rpressure.melt, aes(x=hosts, y=propRegional,fill=variable))+geom_boxplot(outlier.size = 0.2,color="black",lwd=0.2,position = position_dodge(preserve = "single"))+scale_fill_brewer(palette ="BrBG")+ylab("Regional fraction\nof plots affected")+xlab("")+scale_fill_manual(values=rev(p2col)[2:4])

gghostBA <- ggplot(Rpressure.melt, aes(x=hosts, y=hostBA,fill=host))+geom_boxplot(outlier.size = 0.2,color="black",lwd=0.2,position = position_dodge(preserve = "single"))+scale_fill_brewer(palette ="BuPu")+ylab(expression(atop("Regional basal", paste("area (", m^2, ha^-1, ")"))))+xlab("")+theme(axis.text.x = element_blank(), axis.title.x = element_blank())

ggpressure <- ggplot(Rpressure.melt, aes(x=hosts, y=pressure,fill=variable))+geom_boxplot(outlier.size = 0.2,color="black",lwd=0.2,position = position_dodge(preserve = "single"))+scale_fill_manual(values=rev(p2col)[2:4])+ylab("Regional pressure")+xlab("host composition")+theme(axis.text.x = element_text(angle = 45, hjust=1,vjust=1), axis.title.x = element_blank(), legend.position = "top", legend.title = element_blank())

ggcombo <- gghostBA + theme(axis.text.x=element_blank()) + ggpropRegional +  plot_layout(ncol=1)+ plot_annotation(tag_levels = 'A')+ plot_layout(guides = "collect") & theme(legend.position='bottom', legend.title=element_blank(),legend.box="vertical", legend.margin=margin(), plot.tag = element_text(face = 'bold'))
ggsave(here("Results/Figures/Figure-PressureComboByhostIdentity.jpg"), plot=ggcombo , width=fig.width1.5, height=fig.width1*1.5, units="in")

knitr::include_graphics(here("Results/Figures/Figure-PressureComboByhostIdentity.jpg"))
```

```{r FigClimate, fig.cap = "A caption", out.width = "244.8pt"}
ggDecTMINa <- ggplot(datBB.c, aes(x=hostsFPlong, y=DecTMINa))+geom_boxplot(outlier.size = 0.2,color="black",lwd=0.2)+ylab("Dec TMIN\n Anomaly")

ggJulTMAXa<- ggplot(datBB.c, aes(x=hostsFPlong, y=JulTMAXa))+geom_boxplot(outlier.size = 0.2,color="black",lwd=0.2)+ylab("Jul TMAX\n Anomaly")

ggPPTa<- ggplot(datBB.c, aes(x=hostsFPlong, y=PPTa))+geom_boxplot(outlier.size = 0.2,color="black",lwd=0.2)+ylab("Annual PPT\n Anomaly")

ggJJSPEI <- ggplot(datBB.c, aes(x=hostsFPlong, y=JJSPEI))+geom_boxplot(outlier.size = 0.2,color="black",lwd=0.2)+ylab("Jun-Jul SPEI")

ggWYSPEI <- ggplot(datBB.c, aes(x=hostsFPlong, y=WYSPEI))+geom_boxplot(outlier.size = 0.2,color="black",lwd=0.2)+ylab("5-year SPEI")+theme(axis.text.x = element_text(angle = 45, hjust=1,vjust=1), axis.title.x = element_blank())

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = F,
	message = FALSE,
	warning = FALSE,
	dpi = 300,
	progress = FALSE,
	cache = FALSE
)

@@

```{r determine co-occurrence of hosts}
Fig.file <- here("Results/Figures/Figure-AgentbyHostcombo.jpg")
df$hostsFPshort<- df$hostsFPlong
df$hostsFPshort <- factor(df$hostsFPlong, levels=levels(df$hostsFPlong), labels=c("PICO", "PIEN", "ABLA", "PICO\n&\nPIEN", "PICO\n&\nABLA", "PIEN\n&\nABLA", "PICO,\nPIEN,\n&\nABLA"))
df$agentsFplongrev <- df$agentsFplong
df$agentsFplongrev <- factor(df$agentsFplongrev, levels=rev(levels( df$agentsFplong)))
jpeg(Fig.file, width=fig.width2,height=fig.width1.5,units="in",res=300)
ggplot(df[df$focalhost=="Any",], aes(x=hostsFPshort,fill=agentsFplongrev))+geom_bar()+theme(legend.title = element_blank())+xlab("hosts")+scale_fill_manual(values=p2col)
whatever <- dev.off()
```

### Determine suitable hosts
Construct random forest models using synthetic minority oversampling technique (SMOTE)

#### MPB
```{r fit MPB RF model, eval=F}
response <- "agentMPB"
vars <- c("BA.PICO", "QMD.PICO", "propBA.PICO")
datBB.c <- df[df$focalhost=="Any",]

@@

```{r determine suitable stands}
MPBpred <- read.csv(here("Results/RF-pred-agentMPB-occurrence.csv"))
SBpred <- read.csv(here("Results/RF-pred-agentSB-occurrence.csv"))
WBBBpred <- read.csv(here("Results/RF-pred-agentWBBB-occurrence.csv"))

@@

threshold <-0.5
BBpred$MPBsuitable <- ifelse(BBpred$MPBpred > threshold, "MPB", "absent")
BBpred$SBsuitable <- ifelse(BBpred$SBpred > threshold, "SB", "absent")
BBpred$WBBBsuitable <- ifelse(BBpred$WBBBpred >threshold, "WBBB", "absent")

BBpred$agentsuitable <-  do.call(paste, c(BBpred[c("SBsuitable", "MPBsuitable", "WBBBsuitable")],sep="-"))
BBpred$agentsuitable <- factor(BBpred$agentsuitable, levels=c("absent-absent-absent","absent-MPB-absent","SB-absent-absent", "SB-MPB-absent", "absent-absent-WBBB", "SB-absent-WBBB", "absent-MPB-WBBB", "SB-MPB-WBBB"), labels=c("none", "MPB", "SB", "MPB & SB", "WBBB", "SB & WBBB", "MPB & WBBB", "MPB, SB & WBBB"))

BBpred$suitablehosts <- factor(BBpred$agentsuitable, levels=c("none", "MPB", "SB", "WBBB", "MPB & SB",  "MPB & WBBB", "SB & WBBB",  "MPB, SB & WBBB"), labels=c("none", "PICO", "PIEN", "ABLA","PICO & PIEN",  "PICO & ABLA", "PIEN & ABLA",  "PICO, PIEN, & ABLA"), ordered=T)

## Host co-occurrence

### Given suitable stand conditions, is co-occurrence of suitable hosts common?
```{r determine co-occurrence of suitable stands}
bbo.dat<- merge(BBpred, df, id.vars="PLT_CN", all=T)
bbo.dat <- bbo.dat[bbo.dat$focalagent=="Any",]
cocount <-  as.data.frame(table(bbo.dat$agentsFplongrev))
MPBco <-sum(cocount[cocount$Var1 %in% c("MPB & SB",  "MPB & WBBB","MPB, SB & WBBB"),]$Freq)
SBco <-sum(cocount[cocount$Var1 %in% c("MPB & SB",  "SB & WBBB","MPB, SB & WBBB"),]$Freq)
WBBBco <-sum(cocount[cocount$Var1 %in% c("MPB & WBBB",  "SB & WBBB","MPB, SB & WBBB"),]$Freq)

cocount <-  as.data.frame(table(bbo.dat$suitablehosts))
PICOco <-sum(cocount[cocount$Var1 %in% c("PICO & PIEN",  "PICO & ABLA","PICO, PIEN & ABLA"),]$Freq)
PIENco <-sum(cocount[cocount$Var1 %in% c("PICO & PIEN",  "PIEN & ABLA","PICO, PIEN & ABLA"),]$Freq)
ABLAco <-sum(cocount[cocount$Var1 %in% c("PICO & ABLA",  "PIEN & ABLA","PICO, PIEN & ABLA"),]$Freq)

codat <- data.frame(agent =c( "MPB", "SB", "WBBB"), agentco=c(MPBco, SBco, WBBBco), hostco=c(PICOco, PIENco, ABLAco))

bbo.dat$hostsFPshort<- bbo.dat$hostsFPlong
bbo.dat$hostsFPshort <- factor(bbo.dat$hostsFPlong, levels=levels(bbo.dat$hostsFPlong), labels=c("PICO", "PIEN", "ABLA", "PICO\n&\nPIEN", "PICO\n&\nABLA", "PIEN\n&\nABLA", "PICO,\nPIEN,\n&\nABLA"))
bbo.dat$agentsFplongrev <- bbo.dat$agentsFplong
bbo.dat$agentsFplongrev <- factor(bbo.dat$agentsFplongrev, levels=rev(levels( bbo.dat$agentsFplong)))

ggfreq <- ggplot(bbo.dat[bbo.dat$focalhost=="Any",], aes(x=hostsFPshort,fill=agentsFplongrev))+geom_bar()+theme(legend.title = element_blank())+xlab("hosts")+scale_fill_manual(values=p2col)

ggsave( here("Results/Figures/FigureAgentfrequencybyhosts.jpg"), plot=ggfreq,   width=fig.width2,height=fig.width1.5,units="in")

### Is occurrence more likely or severity greater in stands with multiple hosts?

```{r occurrence by suitable hosts}
dfpred <- merge(BBpred, df, id.vars="PLT_CN", all=T)
mpb <- as.data.frame(table(dfpred[dfpred$focalagent=="MPB","affected"], dfpred[dfpred$focalagent=="MPB","suitablehosts"]))
sb <- as.data.frame(table(dfpred[dfpred$focalagent=="SB","affected"], dfpred[dfpred$focalagent=="SB","suitablehosts"]))
wbbb <- as.data.frame(table(dfpred[dfpred$focalagent=="WBBB","affected"], dfpred[dfpred$focalagent=="WBBB","suitablehosts"]))
any<- as.data.frame(table(dfpred[dfpred$focalagent=="Any","affected"], dfpred[dfpred$focalagent=="Any","suitablehosts"]))

results <- NULL
for(j in c("PICO & PIEN",  "PICO & ABLA", "PIEN & ABLA",  "PICO, PIEN, & ABLA")){
  anyj <- any[any$Var2==j,]
  mpbj <- mpb[mpb$Var2==j,]
  sbj <- sb[sb$Var2==j,]
  wbbbj <- wbbb[wbbb$Var2==j,]

@@

  res <- data.frame(suitablehosts=NA, barkbeetle=NA, X=NA, p=NA, prop1=NA, prop2 =NA, lower=NA, upper=NA)
  res$suitablehosts <-j
  res$barkbeetle <- "MPB"
  res$X <- x$statistic
  res$p <- x$p.value
  res$prop1 <- x$estimate[1]
  res$prop2 <- x$estimate[2]
  res$lower <- x$conf.in[1]
  res$upper <- x$conf.in[2]
  x <-NULL
  results <- rbind(results, res)

@@

  res <- data.frame(suitablehosts=NA, barkbeetle=NA, X=NA, p=NA, prop1=NA, prop2 =NA, lower=NA, upper=NA)
  res$suitablehosts <-j
  res$barkbeetle <- "SB"
  res$X <- x$statistic
  res$p <- x$p.value
  res$prop1 <- x$estimate[1]
  res$prop2 <- x$estimate[2]
  res$lower <- x$conf.in[1]
  res$upper <- x$conf.in[2]
  x <-NULL
  results <- rbind(results, res)

@@

  res <- data.frame(suitablehosts=NA, barkbeetle=NA, X=NA, p=NA, prop1=NA, prop2 =NA, lower=NA, upper=NA)
  res$suitablehosts <-j
  res$barkbeetle <- "WBBB"
  res$X <- x$statistic
  res$p <- x$p.value
  res$prop1 <- x$estimate[1]
  res$prop2 <- x$estimate[2]
  res$lower <- x$conf.in[1]
  res$upper <- x$conf.in[2]
  x <-NULL
  results <- rbind(results, res)
}

@@

x <- dfpred[is.na(dfpred$suitablehosts)==FALSE,]
x <- x[x$suitablehosts %in% c("PICO & PIEN",  "PICO & ABLA", "PIEN & ABLA",  "PICO, PIEN, & ABLA"),]
x[x$suitablehosts == "PICO & PIEN" & x$focalagent=="WBBB","affected"]<- 0
x[x$suitablehosts == "PICO & ABLA" & x$focalagent=="SB","affected"]<- 0
x[x$suitablehosts == "PIEN & ABLA" & x$focalagent=="MPB","affected"]<- 0

xx <- x %>% group_by(suitablehosts,focalagent) %>% count(affected)
xxx <- x %>% group_by(suitablehosts) %>% count()

xxxx <- merge(xx, xxx, by="suitablehosts")
xxxx$prop <- xxxx$n.x/xxxx$n.y
xxxx <- xxxx[xxxx$affected=="affected",]
xxxx <- xxxx[,c(1,2,6)]
colnames(xxxx) <- c("suitablehosts", "barkbeetle", "prop")

resultsx <- merge(results, xxxx, id.vars=c("suitablehosts", "barkbeetle"), all=T)
resultsx$barkbeetle <- factor(resultsx$barkbeetle, levels=c("MPB", 'SB', "WBBB", "Any"), ordered=T)
resultsx <- resultsx[is.na(resultsx$prop)==FALSE,]
resultsx$suitablehosts <- factor(resultsx$suitablehosts, levels=c("PICO & PIEN",  "PICO & ABLA", "PIEN & ABLA",  "PICO, PIEN, & ABLA"), labels=c("PICO & PIEN",  "PICO & ABLA", "PIEN & ABLA",  "PICO, PIEN, & ABLA"), ordered=T)
resultsx$pplot <- ifelse(resultsx$p<0.05, "*", "")
resultsx$pplot[is.na(resultsx$pplot) ==TRUE] <- ""

ggoccurrence <- ggplot(resultsx, aes(x=barkbeetle,y=prop, fill=barkbeetle))+geom_bar(stat="identity")+theme(legend.title = element_blank())+geom_text(aes(label=pplot), vjust=-0.25,size=6)+facet_grid(.~suitablehosts, scales="free_x", space="free_x")+xlab("bark beetle species")+scale_fill_manual(values=c(p1col[1:3], gray(0.3)))+theme(legend.position = "none")+ylab("proportion of plots")

ggsave( here("Results/Figures/FigureOccurrencebySuitablehosts.jpg"), plot=ggoccurrence,  width=fig.width2, height=fig.width1,units="in")

```{r severity by suitable hosts}
dfpred <- merge(BBpred, df, id.vars="PLT_CN", all=T)
mpb <- dfpred[dfpred$focalagent=="MPB" & dfpred$suitablehosts %in% c("PICO & PIEN",  "PICO & ABLA",   "PICO, PIEN, & ABLA"), ]
sb <- as.data.frame(table(dfpred[dfpred$focalagent=="SB","affected"], dfpred[dfpred$focalagent=="SB","suitablehosts"]))
wbbb <- as.data.frame(table(dfpred[dfpred$focalagent=="WBBB","affected"], dfpred[dfpred$focalagent=="WBBB","suitablehosts"]))
any<- as.data.frame(table(dfpred[dfpred$focalagent=="Any","affected"], dfpred[dfpred$focalagent=="Any","suitablehosts"]))

dfpred <- dfpred[dfpred$suitablehosts %in% c("PICO & PIEN",  "PICO & ABLA", "PIEN & ABLA",  "PICO, PIEN, & ABLA"), ]

dfpred <- dfpred[dfpred$affected01=="affected",]

results <- NULL
for(j in c("PICO & PIEN",  "PICO & ABLA", "PIEN & ABLA",  "PICO, PIEN, & ABLA")){

  x <- kruskal.test(propTBA~focalagent, data=dfpred[dfpred$suitablehosts ==j  & dfpred$focalagent %in% c('Any', 'MPB'),])

  res <- data.frame(suitablehosts=NA, barkbeetle=NA, X=NA, p=NA)
  res$suitablehosts <-j
  res$barkbeetle <- "MPB"
  res$X <- x$statistic
  res$p <- x$p.value
  x <-NULL
  results <- rbind(results, res)

    x <- kruskal.test(propTBA~focalagent, data=dfpred[dfpred$suitablehosts ==j  & dfpred$focalagent %in% c('Any', 'SB'),])

  res <- data.frame(suitablehosts=NA, barkbeetle=NA, X=NA, p=NA)
  res$suitablehosts <-j
  res$barkbeetle <- "SB"
  res$X <- x$statistic
  res$p <- x$p.value
  x <-NULL
  results <- rbind(results, res)

      x <- kruskal.test(propTBA~focalagent, data=dfpred[dfpred$suitablehosts ==j  & dfpred$focalagent %in% c('Any', 'WBBB'),])

  res <- data.frame(suitablehosts=NA, barkbeetle=NA, X=NA, p=NA)
  res$suitablehosts <-j
  res$barkbeetle <- "WBBB"
  res$X <- x$statistic
  res$p <- x$p.value
  x <-NULL
  results <- rbind(results, res)

}

results$pplot <- ifelse(results$p<0.05, "*", NA)

dfpred$pplot <- NA
for(j in c("PICO & PIEN",  "PICO & ABLA", "PIEN & ABLA",  "PICO, PIEN, & ABLA")){

  x <- dfpred[dfpred$suitablehosts ==j  & dfpred$focalagent == 'MPB',]
  xplt <- x[x$propTBA==max(x$propTBA), "PLT_CN"][1]
  dfpred[dfpred$suitablehosts ==j  & dfpred$focalagent == 'MPB' &dfpred$PLT_CN==xplt, ]$pplot <- results[results$suitablehosts==j & results$barkbeetle=='MPB', "pplot"]

    x <- dfpred[dfpred$suitablehosts ==j  & dfpred$focalagent == 'SB',]
  xplt <- x[x$propTBA==max(x$propTBA), "PLT_CN"][1]
  dfpred[dfpred$suitablehosts ==j  & dfpred$focalagent == 'SB' &dfpred$PLT_CN==xplt, ]$pplot <- results[results$suitablehosts==j & results$barkbeetle=='SB', "pplot"]

    x <- dfpred[dfpred$suitablehosts ==j  & dfpred$focalagent == 'WBBB',]
  xplt <- x[x$propTBA==max(x$propTBA), "PLT_CN"][1]
  dfpred[dfpred$suitablehosts ==j  & dfpred$focalagent == 'WBBB' & dfpred$PLT_CN==xplt, ]$pplot <- results[results$suitablehosts==j & results$barkbeetle=='WBBB', "pplot"]

}

dfpred <- dfpred[!(dfpred$suitablehosts == "PICO & PIEN" & dfpred$focalagent=="WBBB"),]
dfpred <- dfpred[!(dfpred$suitablehosts == "PICO & ABLA" & dfpred$focalagent=="SB"),]
dfpred <- dfpred[!(dfpred$suitablehosts == "PIEN & ABLA" & dfpred$focalagent=="MPB"),]

ggseverity <- ggplot(dfpred, aes(x=focalagent, y=propTBA*100, fill=focalagent, colour=focalagent))+geom_boxplot(notch=T)+facet_grid(.~suitablehosts, scales="free_x", space="free_x")+xlab("bark beetle species")+scale_fill_manual(values=adjust_transparency(c(p1col[1:3], gray(0.3)), 0.5))+scale_colour_manual(values=c(p1col[1:3], gray(0.3)))+theme(legend.position = "none")+ylim(1,100)+geom_text(aes(label=pplot), vjust=-0.25,size=10, col="black")+ylab("Severity (% BA affected)")

ggsave(here("Results/Figures/FigureSevertitybySuitablehosts.jpg"), plot=ggseverity, width=fig.width2,height=fig.width1,units="in")

## Agent co-occurrence

### Given suitable stand conditions, is co-occurrence of multiple agents common?

### What are the effects of multiple agents?
```{r severity by suitable hosts}
dfpred <- merge(BBpred, df, id.vars="PLT_CN", all=T)
dfpred <- dfpred[dfpred$suitablehosts %in% c("PICO & PIEN",  "PICO & ABLA", "PIEN & ABLA",  "PICO, PIEN, & ABLA"), ]

dfpred <- dfpred[dfpred$affected01=="affected",]

dfpred <- dfpred[!(dfpred$suitablehosts=="PICO & PIEN" & dfpred$agentsF %in% c("WBBB", "MPB &\nWBBB", "SB &\nWBBB", "MPB,\nSB, &\nWBBB")),]

dfpred <- dfpred[!(dfpred$suitablehosts=="PICO & ABLA" & dfpred$agentsF %in% c("SB", "MPB\n& SB", "SB &\nWBBB", "MPB,\nSB, &\nWBBB")),]

dfpred <- dfpred[!(dfpred$suitablehosts=="PIEN & ABLA" & dfpred$agentsF %in% c("MPB", "MPB\n& SB", "MPB &\nWBBB", "MPB,\nSB, &\nWBBB")),]

ggmultipelseverity <- ggplot(dfpred[dfpred$focalagent=="Any",], aes(x=agentsF, y=propTBA*100, fill=agentsF))+geom_boxplot()+facet_grid(.~suitablehosts, scales="free_x", space = "free_x")+theme(legend.position="none")

ggsave(here("Results/Figures/FigureSeverityMultiple.jpg"), plot=ggmultipelseverity , width=fig.width2,height=fig.width1,units="in")
```

@@

@@

```{r FigCoOccurrence, fig.cap = "Co-occurrence",  out.width = "504.8pt"}

knitr::include_graphics( here("Results/Figures/FigureAgentfrequencybyhosts.jpg"))
```

@@

### Is occurrence more likely or severity greater in stands with multiple hosts?

```{r FigOccurrence, fig.cap = "Severity",  out.width = "504.8pt"}

knitr::include_graphics( here("Results/Figures/FigureOccurrencebySuitablehosts.jpg"))
```

```{r FigSeverity, fig.cap = "Severity", out.width = "504.8pt"}

knitr::include_graphics( here("Results/Figures/FigureSevertitybySuitablehosts.jpg"))
```

@@

## Supplement
### Randon Forest Modeling
```{r FigPartial,  fig.cap = "A caption", out.width = "504.8pt"}
MPBpp <- read.csv(here("Results/RF-partialdependence-agentMPB-occurrence.csv"))
MPBpp$focalagent <- "MPB"
SBpp <- read.csv(here("Results/RF-partialdependence-agentSB-occurrence.csv"))
SBpp$focalagent <- "SB"
WBBBpp <- read.csv(here("Results/RF-partialdependence-agentWBBB-occurrence.csv"))
WBBBpp$focalagent <- "WBBB"

ppplot <- pBA1  + pQMD1 +theme(axis.title.y=element_blank(),axis.text.y=element_blank()) +ppropBA1 +theme(axis.title.y=element_blank(),axis.text.y=element_blank())+ pBA2 + pQMD2 +theme(axis.text.y=element_blank())  + ppropBA2 +theme(axis.text.y=element_blank()) + plot_layout(ncol=3, heights=c(2,1))

ggsave(here("Results/Figures/PartialDependence.jpg"), plot=ppplot, width=fig.width2,height=fig.width1,units="in")
knitr::include_graphics(here("Results/Figures/PartialDependence.jpg"))

### Comparison of stand structure and composition, climate, and regional bark beetle activity  
```{r FigStandSC,  fig.cap = "A caption", out.width = "322.56pt", eval=F}
datBB.c <-df.sub[!df.sub$focalhost=="Any",]

@@

```{r generate maps, eval=F}
bbo.dat<- merge(BBpred, df, id.vars="PLT_CN", all=T)

df.all <- bbo.dat[bbo.dat$focalhost == "Any", ]
